---
title: "wfprogression"
output: html_document
date: "2024-09-18"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(wfprogression)
library(mapview)
```


## Create bounding box polyon
We will create a bounding box polygon around a fire or other polygon 
it will be buffered to help capture nearby scans or progression polygon
Reproject to GDA94 because the data in the database is GDA94
We will also set the start and end dates for search all image sources

```{r}

#run bounding box function
#"dat.fire.polygon" is an example fire area that comes with the package
dat.bbox <-fire_bbox_polygon(fire_polygon = dat.fire.polygon,
                             bbox_buffer_km = 20,new_crs = 4283)


#inspect bounding box
mapview(dat.bbox)+mapview(dat.fire.polygon,col.regions="red")
```


for all the image search functions we need a start date and end date to search
For all the image download function we need an output folder
```{r}
#get start and end dates (or insert your own)
datestart <- dat.fire.polygon$StartDate
dateend <- "2023-12-08"

#set an output path for all the images that will be downloaded
outdir <- "C:\\temp\\tifs2"
dir.create(outdir)

```





We can also draw a polygon for the area we would like to search
```{r}
dat.aus <- rnaturalearth::ne_countries(country = "Australia") %>% 
  dplyr::select(brk_name)

map <- mapview(dat.aus)
mapedit::drawFeatures(map)

```

## Search for satellite images
There are functions to search for satellite images of a fire area (using the bounding box polygon).
We can search for:Landsat, Sentinel 2, Sentinel 3, VIIRS (including hotspots), MODIS (including hotspots)

## Landsat and Sentinel 2
we can search for these together using the Geoscience Australia STAC browser
There is a function that searches this database, where the inputs are a polygon and dates.
It returns the paths of three bands from each satellite product that can be used to make false colour images
Here we'll get the start and end dates from the fire polygon, which was originally from NSW Fire History

```{r}


#run function. Leave "collection_names" to default to all Landsat and Sentinel 2
dat.stac <- fire_search_stac(fire_bbox = dat.bbox,start_date = datestart,end_date = dateend)
head(dat.stac)

```

Download the data
The images file are cloud-optimised geotiffs, which means we can only download the part of the image we need i.e. crop before downloading
Inputs are the data frame produced by fire_search_stac and an output folder
```{r}

fire_download_stac(dat.bbox,dat.stac,outdir)
```

## Sentinel 3
Sentinel 3 are course scale images, more similar to MODIS.
These are not held by Geoscience Australia, but we can search for and download from the copernicus STAC browser.
Sentinel 3 data may only be useful for large fires where nothing else is available, given its coarse resolution.
collection_names can be changed to search for other available products, see here: https://documentation.dataspace.copernicus.eu/APIs/STAC.html

```{r}
dat.sen3 <- fire_search_copernicus(fire_bbox = dat.bbox,start_date = datestart,end_date = dateend)

```

Download the data
These are not cloud-optimised geotiffs, so we need to download the whole files, which can be slow.
There is then a seperate function to create the false colour RGBs.
We can filter the data frame to download just some of the images.
A copernicus account is required to download the data. I have defaulted the user name and password to my mstorey account, but this would need to be changed. Sign up here:
https://dataspace.copernicus.eu/


```{r}
#downloading these files takes a long time
#but we can extract just some of the files to download e.g. daytime images
dat.sen3.filt <- dat.sen3 %>% 
  dplyr::filter(lubridate::hour(datetimelocal) %in% 8:19)

fire_download_copernicus(fire_bbox = dat.bbox,download_df = dat.sen3.filt,dest_folder = outdir)
```

convert downloaded sentinel 3 zips to geotiffs
Input path is folder where SEN3.zips where written. The function will list and convert all SEN3.zips in that folder to geotiffs.
There are no atmospheric or topographic corrections performed.
The geotiffs will be cropped to fire_bbox
```{r message=FALSE, warning=FALSE}
fire_sentinel3_tif(fire_bbox = dat.fire.polygon,
                    sen3_folder = outdir)

```



## MODIS and VIIRS
I can't find where to easily download raw MODIS and VIIRS images which have image capture timestamps.
Instead this function captures screenshots from the NASA GIBS WMS, which only has date, and overlays hotspots, which have timestamps.
Occasionally more than one set of hotspots are associated with one MODIS image, so both sets are mapped 
This function calls other functions to download hotspots, access the wms map for each date, take a screenshot and create geotiffs.
Different GIBS products are accessed for day (false colour RGB) and night (thermal band)
On days where there were no hotspots, the image is written with only date in the file name
```{r message=FALSE, warning=FALSE}
fire_hotspot_map(fire_bbox = dat.bbox,
                 start_date = datestart,
                 end_date = dateend,
                 dest_folder = outdir)



```


## Himiwari images
BOM keeps an updated collection of 10-min Himiwari cloud optimised geotiffs
These can be accessed via a thredds server.
There is a function to search and return a dataframe of available images.
Set the timestep_minutes argument to find images at selected timesteps, e.g. every 10 minutes or 60 minutes.
We can then filter that and download selected images, cropped to our bounding box polygon or fire

```{r}

dat.himiwari <- fire_search_himawari(fire_bbox = dat.bbox,start_date = datestart, end_date = dateend,timestep_minutes = 120)

```
Download the data
Filter this data frame first to download one day as an example.
Only the part of image within the bbox will be downloaded. 
As a default, only bands used to create a false colour image will be downloaded.
A false colour geotiff is saved for day times and a single band infrared is saved for night times.
```{r}
dat.himiwari.filt <- dat.himiwari %>% 
  #retain only first date and day time images
  dplyr::filter(date_local==min(date_local),daynight=="day")

fire_download_himiwari(fire_bbox = dat.bbox,df_download = dat.himiwari,dest_folder = outdir)
```

```{r}
dat.tifs <- list.files(outdir,pattern="\\.tif$",full.names = T)

purrr::map(dat.tifs,~fire_write_plot(.x,dat.fire.polygon,dat.bbox,outdir))
```

